Sparse Function-space Representation \\ of Neural Networks

### 2023-05-24 paper/main.tex

\section{Method details}
\label{app:method}
%
Key to our method is the ability to compute the dual parameters in \cref{eq:dual_param}. For general exponential family likelihoods we can derive simpler expressions:
\begin{equation}
 \ell(\sigma(f), y) = -\log p(y \mid f) =  A(f)  - \langle y, f\rangle
\end{equation}
where $A(f)$ is the log partition function and $\sigma(.)$ is the link function. Taking the first and second derivative of the loss in this form then become:

@@

\section{\our for continual learning}
\label{app:cl}
This section provides additional details about the derivation of the \our-based regularizer described in \cref{sec:sequential} and how we extended it for dealing with the multi-output setting in our experiments.
\subsection{Derivation of the regularizer term}
For continual learning, the model cannot access the training data from previous tasks, denoted here as $\dataset_\textrm{old} = \bigcup_{s=1}^{t-1} \dataset_s$, and for each new task receives a new chunk of data $\dataset_\textrm{new} = \dataset_t$. On the other hand, in the supervised learning setting (described in \cref{sec:methods}), the neural network model has access to all the training data at once, \ie, the objective we are trying to optimize is
\begin{equation}
	\weights^* = \arg \min_\weights \mathcal{L}(\dataset_\textrm{old} \cup \dataset_\textrm{new}, \weights) = \arg \min_\weights \sum_{i \in \dataset_\textrm{old} \cup \dataset_\textrm{new}}\ell(f_\weights(\mathbf{x}_{i}), y_i) + \delta \mathcal{R}(\weights).
\end{equation} 
Our goal in the continual setting is to replace the missing data coming from $\dataset_\textrm{old}$ by resorting to a functional regularizer that accounts for that by forcing the predictions of the neural networks over the inducing points $\mathcal{Z}_t$ to come with high probability from $f_\weights(\vz_i) \sim q(\vu),  \forall \vz_i \in \mathcal{Z}_t$. As shown in \citep{khan2017conjugate, adam2021dual}, we can view the dual parameters derived in \cref{sec:sparse-dual-gp}, as {\color{red} TODO}. Then, we can rewrite the sparse GP posterior for $q(\vu)$ in terms of the following approximate normal distributions:
%
\begin{align}
	q(\vu)
	& \propto p_\weights(\vu) \Norm(\tilde{\vy}; \vu, \MB_\vu)\\
	& \propto  \Norm(0, \MKzz) \prod_{i \in \dataset_\textrm{old}} \exp \! \left(-\frac{\hat{\beta}_i}{2}(\tilde{y}_i - \va_i^\top \vu)^2 \right) \\
	& \propto \Norm(0, \MKzz) \, \exp\left((\tilde{\vy} - \vu)^\top \bar{\MB}^{-1}(\tilde{\vy} - \vu)\right)
\end{align}
% The first normals are the approximate likelihoods 
where $\va_i^\top = \vkzi^\top \MKzz^{-1}$ and $\tilde{y}_i = \hat{\alpha}_i + f_{\vw^*}(\vx_i)\hat{\beta}_i$.
The sparse GP posterior $q(\vu)$ is proportional to the product between the prior over the weights $p_\weights(\vu)$ and a Gaussian term $\Norm(\tilde{\vy}; \vu, \MB_\vu)$, which arises from {\color{red} TODO}

Replacing $\va_i^\top \vu$ with $\vx_i^{\top}\vw$, we can see the similarity between the sparse GP and a linear model where the kernel matrix determines the prior. To arrive at the MVN Gaussian, we need to rearrange the quadratic terms to match the prior sufficient statistics. Performing conjugate summation of prior and likelihood, we would arrive at the sparse GP posterior at the inducing points. The correlated covariance structure $\MSigma$ and the corresponding $\tilde{\vy}$ are simply a different form of the sparse dual variables,
\begin{equation}
\quad \bar{\MB}^{-1} =  \MKzz \MB_\vu^{-1} \MKzz \quad \text{and} \quad \tilde{\vy} = \MKzz \MB_{\vu}\valpha_{\vu}.
\end{equation}
Written as above, we can see that the spare dual variables can be interpreted as a sparse MVN representation.

@@

\section{\our for model-based reinforcement learning}
\label{app:rl}
%
As a final experiment in the main paper in \cref{sec:rl-exp}, we provided a demonstration of combining our \our approach with model-based reinforcement learning. As such, this is a direct extension of the uncertainty quantification and representation capabilities already demonstrated in the supervised (and continual) learning examples in the main paper. However, the RL setting requires slightly more background to it and thus \cref{app:rl} has been separated out to a separate section in the appendix.

>>>>

\section{Method details}
\label{app:method}
%
\todo{This reads a bit off, is it grammatically correct?} Key to our method is the ability to compute the dual parameters as in \cref{eq:dual_param}. For general exponential family likelihoods we can derive simpler expressions:
\todo{How does this connect to Eq 5?}
\begin{equation}
 \ell(\sigma(f), y) = -\log p(y \mid f) =  A(f)  - \langle y, f\rangle
\end{equation}
where $A(f)$ is the log partition function and $\sigma(.)$ is the link function. Taking the first and second derivative of the loss in this form then becomes: \todo{Becomes what, to be written?}

@@

\section{\our for continual learning}
\label{app:cl}
This section provides additional details about the derivation of the \our-based regularizer described in \cref{sec:sequential} and how we extended it for dealing with the multi-output setting in our experiments.
\subsection{Derivation of the regularizer term}
For continual learning, the model cannot access the training data from previous tasks, denoted here as $\dataset_\textrm{old} = \bigcup_{s=1}^{t-1} \dataset_s$, and for each new task receives a new chunk of data $\dataset_\textrm{new} = \dataset_t$. On the other hand, in the supervised learning setting (described in \cref{sec:methods}), the neural network model has access to all the training data at once, \ie, the objective we are trying to optimize is
\begin{equation}
	\weights^* = \arg \min_\weights \mathcal{L}(\dataset_\textrm{old} \cup \dataset_\textrm{new}, \weights) = \arg \min_\weights \sum_{i \in \dataset_\textrm{old} \cup \dataset_\textrm{new}}\ell(f_\weights(\mathbf{x}_{i}), y_i) + \delta \mathcal{R}(\weights).
\end{equation} 
Our goal in the continual setting is to replace the missing data coming from $\dataset_\textrm{old}$ by resorting to a functional regularizer that accounts for that by forcing the predictions of the neural networks over the inducing points $\mathcal{Z}_t$ to come with high probability from $f_\weights(\vz_i) \sim q(\vu),  \forall \vz_i \in \mathcal{Z}_t$. As shown in \citep{khan2017conjugate, adam2021dual}, we can view the dual parameters derived in \cref{sec:sparse-dual-gp}, as {\color{red} TODO}. Then, we can rewrite the sparse GP posterior for $q(\vu)$ in terms of the following approximate normal distributions:
%
\begin{align}
	q(\vu)
	& \propto p_\weights(\vu) \Norm(\tilde{\vy}; \vu, \MB_\vu)\\
	& \propto  \Norm(0, \MKzz) \prod_{i \in \dataset_\textrm{old}} \exp \! \left(-\frac{\hat{\beta}_i}{2}(\tilde{y}_i - \va_i^\top \vu)^2 \right) \\
	& \propto \Norm(0, \MKzz) \, \exp\left((\tilde{\vy} - \vu)^\top \bar{\MB}^{-1}(\tilde{\vy} - \vu)\right),
\end{align}
% The first normals are the approximate likelihoods 
where $\va_i^\top = \vkzi^\top \MKzz^{-1}$ and $\tilde{y}_i = \hat{\alpha}_i + f_{\vw^*}(\vx_i)\hat{\beta}_i$.
The sparse GP posterior $q(\vu)$ is proportional to the product between the prior over the weights $p_\weights(\vu)$ and a Gaussian term $\Norm(\tilde{\vy}; \vu, \MB_\vu)$, which arises from {\color{red} TODO}

Replacing $\va_i^\top \vu$ with $\vx_i^{\top}\vw$, we can see the similarity between the sparse GP and a linear model where the kernel matrix determines the prior. To arrive at the MVN \todo{What is MVN short for?}Gaussian, we need to rearrange the quadratic terms to match the prior sufficient statistics. When performing conjugate summation of prior and likelihood, we arrive at the sparse GP posterior at the inducing points. The correlated covariance structure $\MSigma$ and the corresponding $\tilde{\vy}$ are simply a different form of the sparse dual variables,
\begin{equation}
\quad \bar{\MB}^{-1} =  \MKzz \MB_\vu^{-1} \MKzz \quad \text{and} \quad \tilde{\vy} = \MKzz \MB_{\vu}\valpha_{\vu}.
\end{equation}
Written as above, we can see that the sparse dual variables can be interpreted as a sparse MVN representation.

@@

\section{\our for model-based reinforcement learning}
\label{app:rl}
%
As a final experiment in the main paper in \cref{sec:rl-exp}, we provided a demonstration of combining our \our approach with model-based reinforcement learning. As such, this is a direct extension of the uncertainty quantification and representation capabilities already demonstrated in the supervised (and continual) learning examples in the main paper. However, as the RL setting differs from the supervised and continual learning settings, we present additional background material on it and thus \cref{app:rl} has been separated out to a separate section in the appendix.
